Variable Selection

### 2015-07-07 cluster.rmd
began clustering

---
title: "Clustering"
author: "Amy Cook"
date: "Thursday, July 02, 2015"
output: word_document
---

@@

Full:
* inv.mlsto
* timespan
* no.users
* Discipline
* pc.pro
* client.totinv
* code.director
* Num.disc
* Business
* Biz.type

Many NA:
* code.contact
* code.client
* Project.Value
* JD.Second
* Billing.Type
* Post.Code 
* code.ProjEng
* Billing.Type
* majority.pos

>>>>

---
title: "clusters"
output: word_document
---

```{r, echo=FALSE, include=FALSE}

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# library('dendextend')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")

all7c<- read.csv('all7c.csv')[,-1]
all7c$Post.Code<- as.factor(all7c$Post.Code)

```

@@

Full:

inv.mlsto
timespan
no.users
Discipline
pc.pro
client.totinv
code.director
Num.disc
Business
Biz.type

Many NA:

code.contact
code.client
Project.Value
JD.Second
Billing.Type
Post.Code
code.ProjEng
Billing.Type
majority.pos

Create reduced dataset and have a think about what you want to cluster against each other

```{r final_var, echo=FALSE}

all7d<- all7c %>% select(inv.mlsto, timespan, no.users, Discipline, pc.pro, client.totinv, code.director, Num.disc, Business,
                         Biz.type, code.contact, pc.majpos,
                         code.client, Project.Value, JD.Second, Billing.Type, Post.Code, code.ProjEng, Billing.Type, majority.pos)

```

What kind of clusters are you looking for? Within the numeric variables:

```{r, echo=FALSE}
all7c[,!sapply(all7c, is.factor)] %>% colnames
all7d[,!sapply(all7d, is.factor)] %>% colnames

```

Numeric variables in reduced dataset are: inv.mlsto, timespan, no.users, pc.pro, client.totinv, Num.disc, Project.Value, pc.majpos

Numeric variables worth considering in full dataset: Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto, cost.mlsto,

So, lets choose:

* inv.mlsto
* timespan
* no.users
* pc.pro
* client.totinv
* Num.disc
* Num.days
* mean.peeps
* hours.perday
* balance.mlsto
* hrs.mlsto
* cost.mlsto

Many NA - do second
* pc.majpos
* Project.Value

Before clustering, need to standardise variables, and have a look at 'normality' of other variables

```{r var_check, echo=FALSE}

#Q-Q plot of each variable
clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, pc.majpos, return.pdol)

#get rid of zeroes for log transform

clust[clust$timespan ==0,]$timespan <- 1
clust[clust$pc.pro ==0,]$pc.pro <- 5
clust[clust$hours.perday ==0,]$hours.perday <- .1
clust[clust$hours.perday ==0,]$hours.perday <- .1
clust[clust$balance.mlsto ==0,]$balance.mlsto <- 1
clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1
clust[clust$return.pdol ==0,]$return.pdol <- .01

for(i in 1:ncol(clust)){
qqPlot(clust[,i],
       main=colnames(clust)[i])
}

#qqplots for log of each variable
for(i in 2:ncol(clust)+1){
qqPlot(clust[,i] %>% log,
       main=colnames(clust)[i])
}

#qqplots for sqrt of each variable
for(i in 2:ncol(clust[,-1])+1){
qqPlot(clust[,i] %>% sqrt,
       main=colnames(clust)[i])
}

#density plots - plain variables
for(i in 1:ncol(clust)){
print(
        ggplot(clust, aes_string(x=colnames(clust[i]))) + geom_histogram(aes(y=..density..))
        )
}
#density plots - log transformed variables
clust.log<- log(clust)
for(i in 1:ncol(clust.log)){
        print(
                ggplot(clust.log, aes_string(x=colnames(clust.log[i]))) + geom_histogram(aes(y=..density..))
                )
}

#plot sqrt of timespan
clust$timespan <- clust$timespan %>% sqrt
ggplot(clust, aes_string(x='timespan')) + geom_histogram(aes(y=..density..))

```

taking the log worked for:
* cost.mlsto, hrs.mlsto, hours.perday, Num.days, client.totinv, no.users, inv.mlsto
* sqrt of timespan

so we end up with a numeric data set with the above variables log transformed.
Then scale each by subtracting the mean of each variable and then dividing by the sd of each variable.

```{r clust final, echo=FALSE}
clust<- all7c %>% select(mlsto, inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol)

#get rid of zeroes for log transform

clust[clust$timespan ==0,]$timespan <- 1
clust[clust$hours.perday ==0,]$hours.perday <- .1
clust[clust$hrs.mlsto ==0,]$hrs.mlsto <- 1
clust[clust$cost.mlsto ==0,]$cost.mlsto <- 1

#take log of all variables to be logged in a for loop
logged<- c('cost.mlsto', 'hrs.mlsto', 'hours.perday', 'Num.days', 'client.totinv',
           'no.users', 'inv.mlsto')
for(i in 1:length(logged)){
        clust[,logged[i]]<- clust[,logged[i]] %>% log
}
clust$timespan<- sqrt(clust$timespan)

#now to standardise
clust<- scale(clust[,-1]) %>% as.data.frame
```

Now to do hierarchical clustering

```{r hierarch, echo=FALSE}

#Ward's clustering

x<- dist(clust, method = 'euclidean')
fit<- hclust(x, method = 'ward.D2')
plot(fit)

rect.hclust(fit, k=5, border='red')
#5 clusters looks pretty good

#try horizontal dend with branches coloured for 8 clusters - very slow
# dend<- fit %>% as.dendrogram
# dend %>% color_branches(k=8) %>% plot(horiz=T)
# rect.dendrogram(dend, k=8, horiz=T)

#how many points in each cluster?
clust$groups<- cutree(fit,k=5)
table(clust$groups)

```

Now look at properties of groups using k means clustering

```{r kmeans, echo=FALSE}

k1<- kmeans(clust %>% select(-groups), 5)
k1.centres<- k1$centers %>% as.data.frame
k1.centres$group<- rownames(k1.centres)
k1.melt<- melt(k1.centres)
k1.melt$group<- k1.melt$group %>% as.factor
ggplot(k1.melt, aes(x=variable, y= value, fill=variable)) + geom_bar(stat='identity', position='dodge')+
        facet_wrap(~group) +
        theme(axis.text.x=element_text(angle=45,hjust=1))

#try plotting means of actual values

all7c$groups<- k1$cluster
kmeans.plot<- all7c %>% select(inv.mlsto, timespan, no.users, pc.pro, client.totinv,
                         Num.disc, Num.days, mean.peeps, hours.perday, balance.mlsto, hrs.mlsto,
                         cost.mlsto, return.pdol, groups)
all7c.plot<- ddply(kmeans.plot, .(groups), numcolwise(mean))
all7c.plot$groups<- rownames(all7c.plot)
k2.melt<- melt(all7c.plot)

#summary means of all numeric variables in 5 clusters
clust.summary<- all7c.plot %>% select(-groups) %>% round(2)
tab<- table(all7c$groups) %>% prop.table %>% as.data.frame
bal<- sum(all7c$balance.mlsto)
hrs<-  sum(all7c$hrs.mlsto)
clust.summary$pc.jobs<- round( tab[,2]*100, 1)
clust.summary$bal.pc<- ddply(all7c, .(groups), summarise, bal.pc= round(sum(balance.mlsto)/bal*100, 1)) %>% 
        select(2)
clust.summary$hrs.pc<- ddply(all7c, .(groups), summarise, hrs.pc= round(sum(hrs.mlsto)/hrs*100, 1)) %>% 
        select(2)

clust.summary[,c(1,12,10,11, 13, 9,2,7,6,3,4,8,5,14:16)]

```

### 2015-07-07 var_select.rmd
began clustering

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')

# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
```

@@

@@

#overlapping NA values for variables with many NA's
# first lets look at Job.Detail.Secondary, Job.Type.Primary, Type, JD.Primary, JD.Second
experiment<- all7c
experiment<- experiment %>% select(code.client, Post.Code, Billing.Type, Job.Source, Job.Detail.Secondary, 
                                   Job.Type.Primary, Type, Project.Value,
                                   no.employees, client.age, contact.count, JD.Primary, JD.Second, dist, code.contact,
                                   client.age, ProjEng.Pos, code.ProjEng, majority.pos )
experiment<- sapply(experiment, function(x) is.na(x))
NA.plot<- melt(experiment)
NA.plot$mlsto<- rep(all7c$mlsto, ncol(experiment))
#reorder levels for ggplot by total number of NA TRUE
# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
NA.plot <- within(NA.plot, 
                         Var2 <- factor(Var2, 
                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
                                                                 decreasing=FALSE))))

@@

vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'pos.tech', 'majority.pos',
            c('dist','Post.Code'),
            c('code.contact','contact.count'), c('Job.Type.Primary','JD.Primary'),
            c('ProjEng.Pos', 'code.ProjEng'))

@@

* code.contact - ranks 4th after no.users, timespan, pc.pro
* code.client (ranks 4th/4th)
* Project.Value (ranks 4th/5th)
* JD.Second - ranks 6/6th - before code.director and Num.disc / 5th
* Billing.Type - ranks 3rd last/4th last but I can't see how this is not important/last
* Post.Code - 4th
* code.ProjEng - 8th
* majority.pos

@@

```{r run anova on all, echo=FALSE}
vars.vec<- c('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'dist','Post.Code',
            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary','ProjEng.Pos', 'code.ProjEng', 'pos.tech',
            'majority.pos')

>>>>

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')
# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
```

@@

#create new variable - pc.majpos - percentage done by majority position
all7b<- merge(all7b, testa2 %>% select(mlsto, value), by='mlsto', all.x=T, all.y=F)
#if majority.pos is unknown, then value must be unknown - NA
all7b$value<- ifelse(is.na(all7b$majority.pos), NA, all7b$value)
colnames(all7b)[names(all7b) %in% 'value']<-'pc.majpos'

@@

#overlapping NA values for variables with many NA's
# first lets look at Job.Detail.Secondary, Job.Type.Primary, Type, JD.Primary, JD.Second
experiment<- all7c
experiment<- experiment %>% select(code.client, Post.Code, Billing.Type, Job.Source, Job.Detail.Secondary, 
                                   Job.Type.Primary, Type, Project.Value,
                                   no.employees, client.age, contact.count, JD.Primary, JD.Second, dist, code.contact,
                                   client.age, ProjEng.Pos, code.ProjEng, majority.pos, pc.majpos )
experiment<- sapply(experiment, function(x) is.na(x))
NA.plot<- melt(experiment)
NA.plot$mlsto<- rep(all7c$mlsto, ncol(experiment))
#reorder levels for ggplot by total number of NA TRUE
# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
NA.plot <- within(NA.plot, 
                         Var2 <- factor(Var2, 
                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
                                                                 decreasing=FALSE))))

@@

vars<- list('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees',
            'code.client',
            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'pos.tech', 
            'majority.pos', 'pc.majpos',
            c('dist','Post.Code'),
            c('code.contact','contact.count'), c('Job.Type.Primary','JD.Primary'),
            c('ProjEng.Pos', 'code.ProjEng'))

@@

* code.contact - ranks 4th after no.users, timespan, pc.pro
* code.client (ranks 4th/4th)
* Project.Value (ranks 4th/5th)
* JD.Second - ranks 6/6th - before code.director and Num.disc / 5th
* Billing.Type - ranks 3rd last/4th last but I can't see how this is not important/last
* Post.Code - 4th
* code.ProjEng - 8th
* majority.pos
* pc.majpos

@@

```{r run anova on all, echo=FALSE}
vars.vec<- c('Job.Source', 'Job.Detail.Secondary', 'Type', 'Project.Value', 'no.employees','code.client',
            'client.age', 'contact.count', 'JD.Second', 'Billing.Type', 'dist','Post.Code',
            'code.contact','contact.count', 'Job.Type.Primary','JD.Primary','ProjEng.Pos', 'code.ProjEng', 'pos.tech',
            'majority.pos', 'pc.majpos')
